---
title: "Sample size calculation based on simulated data"
subtitle: "for the study: The relationship between eating and concentration"
author: "Christoph Bamberg"
output:
  pdf_document: default
  html_document: default
  word_document: default
  
---
# using Superpower package by Lakens and colleagues
# bookdown version of documentation: https://aaroncaldwell.us/SuperpowerBook/


```{r}
library(Superpower)
```

# specifying inputs to _ANOVA_design_ function
need to find values for design, n, mu, sd, r, labelnames
_design_: either tested hungry or satiated and with "hungry_good" or "full_good" manipulation. In a counterbalanced 2x2 between-groups design, no within group levels.

_means_: specifying the means of each group in a vector. The order matters.
Setting the order to
1. hungry / "hungry_good"
2. hungry / "full_good"
3. satiated / "hungry_good"
4. satiated / "full_good"
-> do labelnames accordingly

_labelnames_: syntax: (factor1, factor1_level1, factor1_level2, factor2, factor2_level1, factor2_level2)

_standard deviation_: for now assuming homogenous variance in all groups -> one value. Otherwise vector according to same ordering as mu vector. Since I am using small values for mu (around 1), the standard deviation should also not be too large. A value around 0.5 seems good.

_sample size_: fixed for now, later varied to get a power curve. Specified per group - n=20 means 20*4=80 subjects in total
```{r}
#design:
# "b" for between, "w" for within, "*" to combine
design="2b*2b"

#labelnames:
labelnames=c("hunger_manipulation", "hugnry", "satiated", "expectation", "hungry_good", "full_good")

# correlations
# zero for between subject designs
r=0

# standard deviation:
sd=0.5

# sample size:
n=30
```

## different  approaches to specifying mu

### simply setting mu to arbitrary values
```{r}
#means:
# 2*2= 4 means to be specified
mu_test=c(1,2,3,4)#to see whether I correctly ordered the labels
#mu=c(1,-1,-1,1) #if only congurency between conditions had an effect
mu=c(1,-1,0,2) #if satiated performance is better + congruency 
mu=c(1,0.5,0.8,1.2) #smaller differences for both interventions
```

### a more informed approach
procedure: 
1. define wanted difference, d_hunger, for hunger intervention 
2. specify means for each level of the hunger intervention based on d_hunger
3. define difference between the expectations
4. add the expectation difference on top of the hunger difference
```{r}
#### specifying mus based on effect size
#with d= (mu_fed - mu_hungry)/sd
# (assuming same sd for all)

# fixing the difference between hungry / satiated a prior
d_hunger=0.3

mu_hungry=1
#solving for mu_fed:
mu_full=d_hunger *sd + mu_hungry 


# now adding the effect of expectations
d_expectation=0.15

mu_hungry_hgood=mu_hungry+d_expectation*sd
mu_hungry_fgood=mu_hungry-d_expectation*sd
mu_full_hgood=mu_full-d_expectation*sd
mu_full_fgood=mu_full+d_expectation*sd
mu=c(mu_hungry_hgood,mu_hungry_fgood,mu_full_hgood,mu_full_fgood)
print(mu)
```

# plotting the assumed values for mu by simulating draws from a normal distribution
```{r}
set.seed(345)

N=120 #the required sample size, see below

sim_hungry_hgood<-rnorm(N,mean=mu_hungry_hgood,sd=sd)
sim_hungry_fgood<-rnorm(N,mean=mu_hungry_fgood,sd=sd)
sim_full_hgood<-rnorm(N,mean=mu_full_hgood,sd=sd)
sim_full_fgood<-rnorm(N,mean=mu_full_fgood,sd=sd)
condition<-c(rep(1,length.out=N),rep(2,length.out=N),rep(3,length.out=N),rep(4,length.out=N))
condition<-as.factor(condition)
sim_perf<-c(sim_hungry_hgood,sim_hungry_fgood,sim_full_hgood,sim_full_fgood)
#sim_perf<-as.numeric(sim_perf)
DF<-data.frame(cbind(condition,sim_perf))
#View(DF)


library(ggplot2)
ggplot(data=DF, aes(x=condition , y=sim_perf, fill=factor(condition)))+
  geom_violin(scale="area")+
  stat_summary(fun="mean",
               geom="crossbar") +
  labs(x="condition",y="simulated performance",title="Violin plot for simulated performance split by condition",subtitle="black bars indicate means") + scale_fill_discrete(name="Condition",labels=c('hungry/ "hunger good"', 'hungry / "satiated good"','satiated/ "hungry good"', 'satiated/ "satiated good"'))


# different plot
#plot(density(sim_hungry_hgood),col="red",lwd=3)
#lines(density(sim_hungry_fgood),col="yellow",lwd=3)
#lines(density(sim_full_hgood),col="green",lwd=3)
#lines(density(sim_full_fgood),col="blue",lwd=3)


```

# plotting assumed differences between means
```{r}
x=
plot(x=c(1,2,1,2),y=c(mu_hungry_hgood,mu_hungry_fgood,mu_full_hgood,mu_full_fgood),main="Expected effect of interventions",ylab="group means",xlab="expectation manipulation",xaxt="n",yaxt="n")
# Changing x axis
xtick<-c("hungry is good","satiated is good")
axis(1, at=1:2, labels = xtick)
lines(x=c(1,2),y=c(mu_hungry_hgood,mu_hungry_fgood),lwd=3,col="blue")
lines(x=c(1,2),y=c(mu_full_hgood,mu_full_fgood),lwd=3,col="orange") 
legend(1,1.22,legend=c("hungry","satiated"),lwd=3,col=c("blue","orange"))
```

plotting with better visibility
```{r}
x=
plot(x=c(1,2,1,2),y=c(1.025, 0.925, 1.075, 1.225),
    xlim=c(0.95,2.05),ylim=c(0.9,1.25),        
     xaxt="n",yaxt="n",type="n",
    #main="Expected effect of intervention",
    ylab=" Performance",xlab="Expectation manipulation")
# Changing x axis
xtick<-c('"Hungry is good"','"Satiated is good"')
axis(1, at=1:2, labels = xtick)
lines(x=c(1,2),y=c(1.025,mu_hungry_fgood),lwd=3,col="blue")
lines(x=c(1,2),y=c(mu_full_hgood,mu_full_fgood),lwd=3,col="orange") 
legend(0.95,1.23,title="Actual hunger:",legend=c("Satiated","Hungry"),lwd=3,col=c("orange","blue"))
```


# entering the values into design function:
```{r}
design_result<-ANOVA_design(design = design,
                   n = n,
                   mu = mu,
                   sd = sd,
                   labelnames = labelnames)

# to check that design is correctly specified:
#plot(design_result)

simulation_result <- ANOVA_power(design_result,
                                 alpha_level = 0.05,
                                 nsims = 1e3,
                                 verbose = FALSE)

plot(simulation_result$plot1,main="p_value distribution ANOVA")
#plot(simulation_result$plot2,main="p_value distribution paired comparisons")

```

# looking at different sample sizes:
```{r}
plot_power(design_result, max_n = 200,alpha_level = 0.05,desired_power=90)
```




